Prefetching SPARQL query cacher

### 2016-06-01 dissertation/related.tex
profitable query patterns more on the use on a mediator

The query planner first examines all connected subgraphs of a full
SPARQL query, and generates canonical labels for each one. Since it
insists on caching connected subgraphs, it avoids the problem I
encountered in Chapter~\ref{sec:tpfcacheplanning} that the cache may
break up connected subgraphs and thus create Cartesian joins.

The canonical labelling algorithm presented in this paper provides an
opportunity for further work, when considered in an HTTP context: The
\httph{ETag} header, defined in RFC7232 \cite{rfc7232}, can optionally
contain an opaque identifier for a specific version of a resource
representation, e.g. a SPARQL result. Such a canonical label could be
used in conjunction with an algorithm that tracks changes to the RDF
graph to produce an \httph{ETag}. For slow-changing Linked Data, which
is a common case, it is a possibility that a canonical label in
combination with a time-stamp of the entire RDF Graph could yield
considerable benefits. However, the presented algorithm does not cover
SPARQL in its entirety, for example, it does not handle named graphs,
solution modifiers and filters, nor is it clear how it handles
property paths (which in many cases should be trivial, since a
property path often has an equivalent Basic Graph Pattern). Union
queries or negation is also not mentioned, so more work is needed for
this method to have general applicability.

>>>>

The query planner first examines all connected subgraphs of a full
SPARQL query, and generates canonical labels for each one. Since it
insists on caching connected subgraphs, it avoids the problem I
encountered in Chapter~\ref{sec:tpfcacheplanning} that the cache may
break up connected subgraphs and thus create Cartesian joins. To
further enhance the query planner, the cache can optionally include
indexing. The authors also introduce the concept of
\textit{profitable} query patterns. It is not quite clear from the
text, but it seems this implies that the cache can also prefetch
results into the cache, if it finds that the result may benefit
subsequent queries.

The canonical labelling algorithm presented in this paper provides an
opportunity for further work, when considered in an HTTP context: The
\httph{ETag} header, defined in RFC7232 \cite{rfc7232}, can optionally
contain an opaque identifier for a specific version of a resource
representation, e.g. a SPARQL result. Such a canonical label could be
used in conjunction with an algorithm that tracks changes to the RDF
graph to produce an \httph{ETag}. For slow-changing Linked Data, which
is a common case, it is a possibility that a canonical label in
combination with a time-stamp of the entire RDF Graph could yield
considerable benefits. However, the presented algorithm does not cover
SPARQL in its entirety, for example, it does not handle named graphs,
solution modifiers and filters, nor is it clear how it handles
property paths (which in many cases should be trivial, since a
property path often has an equivalent Basic Graph Pattern). Union
queries or negation is also not mentioned, so more work is needed for
this method to have general applicability. Nevertheless, if the entire
Group Graph Pattern of a certain query is supported, and the query
projects all variables, e.g. it wraps the pattern with \sparql{SELECT
  *} or \sparql{CONSTRUCT}, the \httph{ETag} would be trivial to
construct using the presented algorithm in a non-trivial number of
beneficial cases. With this, cached result verification could be done
by a mediator using HTTP requests with the \httph{If-None-Match}
header and so make it possible to use parts of the engine of this
paper usable in the Internet infrastructure.
